Simultaneous Elicitation of Scoring Rule and Agent Preferences for Robust Winner Determination

### 2020-10-10 minimax.tex
Various minor; restored previous 2 phase.

\section{Empirical Evaluation} 
\label{sec:experiments}
%TODO
%\begin{enumerate}
%	\item Reach regret: $n/10$. For this, we need $m^x n$ questions. Wait for more runs (Correct formula)
%Graphs: order symbols in the legend. Reduce number of marks on the X axis.
%MMR and sd with systematically one decimal exactly
%\end{enumerate}
We  performed a number of numerical experiments in order to validate our approach and to test the performance of the different elicitation strategies. %Our experiments are described by topic here below. 
We start by describing the general protocol of experimentation. %use the following protocol, deviations are indicated below for each experiment.
Having picked a problem size $(m, n)$, a number of questions $k$ and a strategy to test, we randomly generate an “oracle”, containing the true preferences of the agents (i.e. the linear orders) and the weights associated with the chair's scoring rule. 
The preferences are generated following an impartial culture assumption: the linear order of each agent is generated independently from the other ones and all have the same probability of appearing. 
For gerenerating the weights we first draw $m-1$ numbers uniformly at random (representing weight ``differences''), normalize and sort them; 
a sequence of convex decreasing weight is then obtained by a decumulative sum.% to the former numbers.
% ordering them, in order to ensure convexity.
The penalty parameters for the Pessimistic and Extended pessimistic strategies are set to $\epsilon_{\text{chair}} = 1.1$ and $\epsilon'_{\text{chair}} = 10^{-6}$ and $\epsilon_{\text{agent}} = 1.0$ and $\epsilon'_{\text{agent}} = 0$.

We start with empty knowledge ($\pprofile = (\emptyset, \emptyset, …), W = \W$) about the preference orderings of the agents or the weight differences favored by the chair. We obtain the first question to be asked using the strategy under test, as described above. We then use the oracle to answer the question and update the system's knowledge, which is thus used to obtain the next question. This is repeated until $k$ answers have been obtained, computinh the resulting MMR values along the way for various values of $k$. We repeat this whole experiment a variable number of times indicated below, for a given $(m, n, k)$, and report the average resulting MMR and the standard deviation \textit{sd}.

%\subsection{Relative evaluation of Pessimistic strategy}
\paragraph{Comparison of strategies in problem instances of small size}

\begin{figure}
	\caption{Average MMR in problems of size $(5, 10)$ after $k$ questions.}
	\label{fig:smallSize}
	\includegraphics[width=.45\textwidth]{comparison.png}
\end{figure}
\begin{table}
	\caption{Average MMR in problems of size $(10, 20)$ after $k$ questions.}
	\label{tab:biggerSize}
	\begin{tabular}{S[table-figures-integer=3, table-number-alignment = right, table-figures-decimal=0]S[table-number-alignment = right]@{ ± }S[table-number-alignment = left]S[table-number-alignment = right]@{ ± }S[table-number-alignment = left]S[table-number-alignment = right]@{ ± }S[table-number-alignment = left]S[table-number-alignment = right]@{ ± }S[table-number-alignment = left]}
		\toprule
		{k} & {Rnd.} & {sd} & {Pes.} & {sd} & {Eli.} & {sd} \\
		\midrule
		0 &20.0 & 0.0 & 20.0 & 0.0&  20.0 & 0.0 \\
		50 & 19.82 & 0.04 &	15.59 &0.56	&17.38& 0.7\\
		100 & 19.12	& 0.26&	11.91&	1.26 &15.23& 0.73\\
		150 & 18.06	& 0.47&	8.92&	1.53 &13.38& 1.11\\
		200 & 16.89	& 0.62&6.24&2.02&11.76& 1.03\\
		250 & 15.63 & 0.84&	4.64&1.72&10.56& 0.89\\
		300 & 14.72	& 0.94&	3.01& 1.35 &10.48& 0.87\\
		350 & 13.41	& 1.15& 1.76& 0.95&10.48& 0.86\\
		400 & 11.98	& 0.68&	0.62&0.56 &10.48& 0.86\\
		\bottomrule
	\end{tabular}
\end{table}
\begin{table}
	\caption{Average MMR in problems of size $(10, 20)$ and geometric weights after $k$ questions.}
	\label{tab:geometricWeights}
	\begin{tabular}{S[table-figures-integer=3, table-number-alignment = right, table-figures-decimal=0]S[table-number-alignment = right]@{ ± }S[table-number-alignment = left]S[table-number-alignment = right]@{ ± }S[table-number-alignment = left]S[table-number-alignment = right]@{ ± }S[table-number-alignment = left]}
		\toprule
		{k} & {Pes.} & {sd} & {Eli.} & {sd} \\
		\midrule
		0	&20.0	&0.0	&20.0	&0.0\\
		100	&12.48	&0.93	&15.6	&0.43\\
		200	&7.43	&1.25	&10.95	&1.12\\
		300	&3.47	&1.32	&6.57	&0.79\\
		\bottomrule
	\end{tabular}
\end{table}
%Our first experimental result concerns small size situations. 
In the first experiment we picked $m = 5, n = 10$ to illustrate our results (variations around this size yield similar conclusions), ran it $200$ times and averaged the results.
\Cref{fig:smallSize} compares some of our strategies in this case.
We see that that Pessimistic performs better than Random,  Extended pessimistic, %. % on a small size problem.
and as well Elitist.
%(The figure also displays the performance of the Elitist strategy).
Asking random questions does not allow to reach a low regret level even after having asked 100 questions, whereas a low regret level ($\MMR$ = 1) is reached by Pessimistic before having asked 60 questions.
We also see that Pessimistic performs slightly better than Extended pessimistic, showing that Pessimistic chooses candidate questions wisely; this is good news since (for $m = 5$ and $n = 10$), Pessimistic is much faster: it takes on average only $24$s for a complete elicitation session ($150$ questions), while Extended pessimistic takes $76$s. Notice that although they exhibit a close performance, Pessimistic performs %significantly %better, in the sense that the numbers are 
consistently better Pessimistic in the different runs of the experiment).
\Cref{tab:biggerSize} shows that these conclusions also hold for a problem of larger size ($m = 10, n = 20$), except that this table does not feature Extended pessimistic, which is too slow to be ran repeatedly on such instance.

@@

%\subsection{Absolute evaluation of Pessimistic strategy}
\paragraph{In-dept evaluation of Pessimistic}
\label{sec:lowRegret}
\begin{table}
	\caption{Number of questions needed by Pessimistic strategy to reach an MMR of $\frac{n}{10}$ (represented by $k_{\MMR}$), by size.}
	\label{tab:lowRegret}
	\begin{tabular}{S[table-figures-integer=1, table-number-alignment = center, table-figures-decimal=0]S[table-figures-integer=1, table-number-alignment = center, table-figures-decimal=0]S[table-number-alignment = center]S[table-figures-integer=1, table-number-alignment = center, table-figures-decimal=0]}
		\toprule
		{m} & {n} & {$\frac{n}{10}$} & {$k_{\MMR}$} \\
		\midrule
		5 & 10 & 1.0 & 60 \\
		5 & 15 & 1.5 & 80 \\
		5 & 20 & 2.0 & 110 \\
		10 & 20 & 2.0 & 340 \\
		10 & 30 & 3.0 & 527 \\
		15 & 30 & 3.0 & 812 \\
		\bottomrule
	\end{tabular}
\end{table}
\begin{table}
	\caption{Number of questions needed by the Pessimistic strategy to elicit the complete preference of one agent.}
	\label{tab:fullProfile}
	\begin{tabular}{S[table-figures-integer=2, table-number-alignment = center, table-figures-decimal=0]S[table-figures-integer=1, table-number-alignment = center, table-figures-decimal=2]}
		\toprule
		m & k \\
		\midrule
		5 & 6.6\\
		10 & 24.8\\
		15 & 48.4\\
		\bottomrule
	\end{tabular}
\end{table}
While in previous experiments we compared Pessimistic to the other strategies, in this set of experiments, we focus %about the absolute quality of 
on Pessimistic. 
We %For practical usability,
think that it is important to know how many questions should be asked in order to reach a low regret level. Here, we consider a low level to be a tenth of the number of agents ($n/10$): such a regret is equivalent to the difference of score of an alternative $x$ that results from switching from a profile $P$ to a profile $P'$ where ten percent of the agents rank $x$ last instead of first.
The results for various sizes are displayed in \cref{tab:lowRegret}. 
We see that the number of questions required to reach a low regret level grows approximately linearly with the number of agents and grows less fast than the square of the number of alternatives, for the sizes we considered. Averaging these data by category of value for $m$, and assuming no questions get asked to the chair, in order to to reach a low regret level, each agent will have to answer $5.6$ questions for $m = 5$, $17.2$ questions for $m = 10$ and $27.1$ questions for $m = 15$. These numbers have to be contrasted with those displayed in \cref{tab:fullProfile}, that estimate the number of questions required to elicit the full preference of an agent. To obtain the data for this table, we constrained the Pessimistic strategy to ask questions to the agents only, and ran it on a single agent problem, with various values for $m$.

@@

We next wonder how fast the regret decreases, measured as a function of the number of questions asked. This is important practically because it may be impossible or undesiderable to ask all questions necessary to reach low level of regrets, as computed in \cref{sec:lowRegret}. Knowing the speed of decrease permits to know which trade-offs can be achieved between the quality of the resulting recommendation and the number of questions that participants have to answer. \Cref{fig:linearity} shows graphically the decrease in MMR according to the number of questions asked for various problem sizes. We see that this function is convex, apart from the very first few questions, indicating %to users of our approach 
that, on average, halving the number of questions will reduce the reduction in regret by less than half.% do less harm than halving the gain in regret, on average.

@@

%In this work, we advocate obtaining an approximate winner without determining completely the voting rule, or the profile. \Cref{sec:lowRegret} evaluates the number of questions per agent required to reach a low regret value.
%In this section we want to explore to which extent it is possible to recommend without knowing the voting rule. One extreme, in our setting, consists in knowing only that the weights are convex, our basis hypothesis. The

%contrast our approach to one that would obtain the full profile on the one hand, and determine the voting rule on the other hand. More precisely, we want to estimate the number of questions that would be required to elicit the complete profile. We focus on the profile because eliciting the weights to full precision would require infinitely many questions, with the questions we use. 
%Note that this may underestimate the number of questions required to determine completely the profile and the rule, as questions to the agents yield no information about the scoring vector.

%To estimate this number, we draw one linear order randomly (with equiprobability over all possible linear orders) and assume the following querying process: given a partial ordering representing our knowledge so far of an agent preferences, we pick a question equiprobably among the remaining edges, and add the edge in the appropriate direction to our partial ordering. We count the average number of questions required to reach a complete ordering.
The experiments so far considered strategies that, at each step, mix questions to the chair and to the agents. One may wonder what is lost in terms of regret by focusing on the Two Phase strategy that asks a bulk of question of one type first and then switches to the other type.  %different proportions of questions to the chair and the agents. Such restrictions may be useful because of (partial) unavailability of the chair, or because the estimated cognitive costs may differ sensibly, as these types of questions differ in nature. 
Notice that there might be practical reasons (for example, the chair may be unavailable during part of the process) for restricting a question type to a specific part of the elicitation process.
We are also interested in analyzing the performance of different proportions of question types, since  (notice that the estimated cognitive costs may differ sensibly, as these types of questions differ in nature).
\Cref{tab:twoP400} shows the MMR value reached in a problem of size $m = 10, n = 20$ after $400$ questions, using the Two phases strategy, in the “ca” ({\em chair then agents}) and in the “ac” ({\em agents then chair}) variants, for different values of $q_{c}$, the number of questions to the chair; 
e.g. the former variant first asks $q_{c}$ questions to the chair and then $400-q_{c}$ to the agents. %, while the latter asks $400-q_{c}$ to the agents and then $q_{c}$ questions to the chair.
These numbers are to be compared with the MMR value reached after 400 questions with the Pessimistic strategy (displayed in \cref{fig:linearity}), which is $0.62$.% ± $0.56$.

These numbers highlight that it is possible to obtain a good-quality recommendation while knowing only that the voting rule is a scoring rule and that the weights are convex, which is our basic hypothesis (corresponding to the line $q_c = 0$, where no question is asked to the chair). By contrast, the Pessimistic strategy (that interleaves the two types of questions) asks on average $48$ questions to the chair.

>>>>

\section{Empirical Evaluation} 
\label{sec:experiments}
%TODO
%\begin{enumerate}
%	\item Reach regret: $n/10$. For this, we need $m^x n$ questions. Wait for more runs (Correct formula)
%Graphs: order symbols in the legend. Reduce number of marks on the X axis.
%MMR and sd with systematically one decimal exactly
%$150$ questions?
%Pessimistic systematically better than XPess?
%\end{enumerate}
We  performed a number of numerical experiments in order to validate our approach and to test the performance of the different elicitation strategies. %Our experiments are described by topic here below. 
We start by describing the general protocol of experimentation. %use the following protocol, deviations are indicated below for each experiment.
Having picked a problem size $(m, n)$, a number of questions $k$ and a strategy to test, we randomly generate an “oracle”, containing the true preferences of the agents (i.e. the linear orders) and the weights associated with the chair's scoring rule. 
The preferences are generated following an impartial culture assumption: the linear order of each agent is generated independently from the other ones and all have the same probability of appearing. 
For generating the weights we first draw $m - 1$ numbers uniformly at random (representing weight ``differences''), normalize and sort them; 
a sequence of convex decreasing weight is then obtained by a decumulative sum.
The penalty parameters for the Pessimistic and Extended pessimistic strategies are set to $\epsilon_{\text{chair}} = 1.1$, $\epsilon'_{\text{chair}} = 10^{-6}$, $\epsilon_{\text{agent}} = 1.0$ and $\epsilon'_{\text{agent}} = 0$.

We start with empty knowledge ($\pprofile = (\emptyset, \emptyset, …), W = \W$) about the preference orderings of the agents or the weight differences favored by the chair. We obtain the first question to be asked using the strategy under test, as described above. We then use the oracle to answer the question and update the system's knowledge, which is thus used to obtain the next question. This is repeated until $k$ answers have been obtained, computing the resulting MMR values along the way for various values of $k$. We repeat this whole experiment a variable number of times indicated below, for a given $(m, n, k)$, and report the average resulting MMR and the standard deviation \textit{sd}.

\paragraph{Comparison of strategies}
\begin{figure}
	\caption{Average MMR in problems of size $(5, 10)$ after $k$ questions.}
	\label{fig:smallSize}
	\includegraphics[width=.45\textwidth]{comparison.png}
\end{figure}
\begin{table}
	\caption{Average MMR in problems of size $(10, 20)$ after $k$ questions.}
	\label{tab:biggerSize}
	\begin{tabular}{S[table-figures-integer=3, table-number-alignment = right, table-figures-decimal=0]S[table-number-alignment = right]@{ ± }S[table-number-alignment = left]S[table-number-alignment = right]@{ ± }S[table-number-alignment = left]S[table-number-alignment = right]@{ ± }S[table-number-alignment = left]S[table-number-alignment = right]@{ ± }S[table-number-alignment = left]}
		\toprule
		{k} & {Rnd.} & {sd} & {Pes.} & {sd} & {Eli.} & {sd} \\
		\midrule
		0 &20.0 & 0.0 & 20.0 & 0.0&  20.0 & 0.0 \\
		50 & 19.82 & 0.04 &	15.59 &0.56	&17.38& 0.7\\
		100 & 19.12	& 0.26&	11.91&	1.26 &15.23& 0.73\\
		150 & 18.06	& 0.47&	8.92&	1.53 &13.38& 1.11\\
		200 & 16.89	& 0.62&6.24&2.02&11.76& 1.03\\
		250 & 15.63 & 0.84&	4.64&1.72&10.56& 0.89\\
		300 & 14.72	& 0.94&	3.01& 1.35 &10.48& 0.87\\
		350 & 13.41	& 1.15& 1.76& 0.95&10.48& 0.86\\
		400 & 11.98	& 0.68&	0.62&0.56 &10.48& 0.86\\
		\bottomrule
	\end{tabular}
\end{table}
\begin{table}
	\caption{Average MMR in problems of size $(10, 20)$ and geometric weights after $k$ questions.}
	\label{tab:geometricWeights}
	\begin{tabular}{S[table-figures-integer=3, table-number-alignment = right, table-figures-decimal=0]S[table-number-alignment = right]@{ ± }S[table-number-alignment = left]S[table-number-alignment = right]@{ ± }S[table-number-alignment = left]S[table-number-alignment = right]@{ ± }S[table-number-alignment = left]}
		\toprule
		{k} & {Pes.} & {sd} & {Eli.} & {sd} \\
		\midrule
		0	&20.0	&0.0	&20.0	&0.0\\
		100	&12.48	&0.93	&15.6	&0.43\\
		200	&7.43	&1.25	&10.95	&1.12\\
		300	&3.47	&1.32	&6.57	&0.79\\
		\bottomrule
	\end{tabular}
\end{table}
Our first experiment concerns small size situations.
\Cref{fig:smallSize} compares some of our strategies in the case $m = 5, n = 10$ (variations around this size yield similar conclusions), where the results are averaged over $200$ runs.
%(The figure also displays the performance of the Elitist strategy, to which we will come back.)
We see that asking random questions does not allow to reach a low regret level even after having asked 100 questions, whereas a low regret level ($\MMR$ = 1) is reached by Pessimistic before having asked 60 questions.
We also see that Pessimistic performs slightly better than Extended pessimistic, showing that Pessimistic chooses candidate questions wisely; this is good news since Pessimistic is much faster: it takes on average only $24$s for a complete elicitation session (for $m = 5$, $n = 10$ and $150$ questions), while Extended pessimistic takes $76$s. Notice that although they exhibit a close performance, Pessimistic performs %significantly %better, in the sense that the numbers are 
consistently better than Extended pessimistic in the different runs of the experiment).
\Cref{tab:biggerSize} shows that these conclusions also hold for a problem of larger size ($m = 10, n = 20$), except that this table does not feature Extended pessimistic, which is too slow to be ran repeatedly on such instance.

@@

%\subsection{Absolute evaluation of Pessimistic strategy}
\paragraph{In-dept evaluation of Pessimistic}
\label{sec:lowRegret}
\begin{table}
	\caption{Number of questions needed by Pessimistic strategy to reach an MMR of $\frac{n}{10}$ (represented by $k_{\MMR}$), by size.}
	\label{tab:lowRegret}
	\begin{tabular}{S[table-figures-integer=1, table-number-alignment = center, table-figures-decimal=0]S[table-figures-integer=1, table-number-alignment = center, table-figures-decimal=0]S[table-number-alignment = center]S[table-figures-integer=1, table-number-alignment = center, table-figures-decimal=0]}
		\toprule
		{m} & {n} & {$\frac{n}{10}$} & {$k_{\MMR}$} \\
		\midrule
		5 & 10 & 1.0 & 60 \\
		5 & 15 & 1.5 & 80 \\
		5 & 20 & 2.0 & 110 \\
		10 & 20 & 2.0 & 340 \\
		10 & 30 & 3.0 & 527 \\
		15 & 30 & 3.0 & 812 \\
		\bottomrule
	\end{tabular}
\end{table}
\begin{table}
	\caption{Number of questions needed by the Pessimistic strategy to elicit the complete preference of one agent.}
	\label{tab:fullProfile}
	\begin{tabular}{S[table-figures-integer=2, table-number-alignment = center, table-figures-decimal=0]S[table-figures-integer=1, table-number-alignment = center, table-figures-decimal=2]}
		\toprule
		m & k \\
		\midrule
		5 & 6.6\\
		10 & 24.8\\
		15 & 48.4\\
		\bottomrule
	\end{tabular}
\end{table}
While in previous experiments we compared Pessimistic to the other strategies, in this set of experiments, we evaluate the Pessimistic strategy in absolute terms. 
We first wonder how many questions should be asked in order to reach a low regret level. Here, we consider a low level to be a tenth of the number of agents ($n/10$): such a regret is equivalent to the difference of score of an alternative $x$ that results from switching from a profile $P$ to a profile $P'$ where ten percent of the agents rank $x$ last instead of first.
The results for various sizes are displayed in \cref{tab:lowRegret}. 
We see that the number of questions required to reach a low regret level grows approximately linearly with the number of agents and grows less fast than the square of the number of alternatives, for the sizes we considered. Averaging these data by category of value for $m$, and assuming no questions get asked to the chair, in order to to reach a low regret level, each agent will have to answer $5.6$ questions for $m = 5$, $17.2$ questions for $m = 10$ and $27.1$ questions for $m = 15$. These numbers have to be contrasted with those displayed in \cref{tab:fullProfile}, that estimate the number of questions required to elicit the full preference of an agent. To obtain the data for this table, we constrained the Pessimistic strategy to ask questions to the agents only, and ran it on a single agent problem, with various values for $m$.

@@

We next wonder how fast the regret decreases, measured as a function of the number of questions asked. This is important practically because it may be impossible or undesirable to ask sufficient questions to reach a low regret level. Knowing the speed of decrease permits to know which trade-offs can be achieved between the quality of the resulting recommendation and the number of questions that participants have to answer. \Cref{fig:linearity} shows graphically the decrease in MMR according to the number of questions asked for various problem sizes. We see that this function is convex, apart from the very first few questions, indicating %to users of our approach 
that, on average, halving the number of questions will reduce the reduction in regret by less than half.% do less harm than halving the gain in regret, on average.

@@

The experiments so far let the strategy target its questions unrestrictively, leaving open the possibility of questioning either the chair or an agent at each step. One may wonder what is lost in terms of regret by asking different proportions of questions to the chair and the agents. Such restrictions may be useful because of (partial) unavailability of the chair, or because the estimated cognitive costs may differ sensibly, as these types of questions differ in nature. \Cref{tab:twoP400} shows the MMR value reached in a problem of size $m = 10, n = 20$ after $400$ questions, using the Two phases strategy, in the “ca” (chair then agents) and in the “ac” (agents then chair) variants. These numbers are to be compared with the MMR value reached after 400 questions with the Pessimistic strategy (displayed in \cref{fig:linearity}), which is $0.62$.% ± $0.56$.
The Pessimistic strategy asks on average $48$ questions to the chair in this setting.

These numbers highlight that it is possible to obtain a good-quality recommendation while knowing only that the voting rule is a scoring rule and that the weights are convex, which is our basic hypothesis (corresponding to the line $q_c = 0$, where no question is asked to the chair).
